### 2013-02-25 web/12.sp.reg/exams/exam3.tex
Spring 2012 reg exam stuff.

\documentclass[11pt]{article}

%% ------------------------------------------------------------------
%% SOLUTIONS:
\def\thel{\noindent\rule{2.5cm}{.5pt}}
\long\def\begsol#1 #2\endsol{#1}
\def\begsol#1{\thel {\bf Solution} \thel}\def\endsol{\relax}
%%uncomment the above line to get the solutions printed

\newcommand\code[1]{\texttt{#1}}
\newcommand\bcode[1]{\texttt{\textbf{#1}}}

%% PROBLEMS:
\def\pts#1{\marginpar{\footnotesize \raggedright  \fbox{#1 {\sc Points}}}}
\newcounter{Pctr}
\newenvironment{problem}{\stepcounter{Pctr}%
\begin{description}
\item[\noindent{\bf Problem} \arabic{Pctr}]
\end{description}}{\relax}
%% ------------------------------------------------------------------

%% SUBPARTS:
\newcounter{parts}
\renewcommand{\theparts}{\Alph}
%% ------------------------------------------------------------------

\begin{document}

\renewcommand{\theenumi}{\Alph{enumi}}
\setcounter{Pctr}{0}

%% ------------------------------------------------------------------

\vfill
\centerline{\Large CS 2510 Exam 3 -- Spring 2012}

~\\[2cm]

\begin{center}
\begin{tabular}{l@{\qquad}l}
Name:                        & \rule{200pt}{.1pt} \\[.5cm]
Student Id (last 4 digits):  & \rule{200pt}{.1pt} \\[.5cm]
\end{tabular}
\end{center}

\noindent\begin{minipage}{7.5cm} $\bullet$ Write down answers in the
space provided.

$\bullet$ You may use all syntax you know from class (that is, the
parts of Java we have studied in class), and you may use any Java
Collection Framework classes you know how to use.

$\bullet$ When defining methods, you do not need to give a complete
class definition---just indicate in which class your method definition
should be placed.

$\bullet$ For tests you only need to provide the expression that
computes the actual value, connecting it with an arrow to the expected
value. For example \code{s.method() -> true} is sufficient.

$\bullet$ Remember that the phrase ``design a class'' or ``design a
method'' means more than just providing a definition. It means to
design them according to the \textbf{design recipe}.  You are
\textit{not} required to provide a method template unless the problem
specifically asks for one.  However, be prepared to struggle if you
choose to skip the template step.

$\bullet$ A word of advice: Decompose complex problems into small
steps.

$\bullet$ We will not answer \textit{any} questions during the exam.

\bigskip

\textit{Good luck!}
\end{minipage}\hfil\begin{minipage}[t]{4.5cm}
\begin{tabular}{|c|l|@/r|}
\hline
\textbf{Problem} & Points & \\ \hline
A & & 10\\ \hline
B & & 15\\ \hline
C & & 10\\ \hline
\textbf{Total} &  & 30 \\ \hline
\end{tabular}
\end{minipage}

\vfill\thispagestyle{empty}
\newpage

%% -----------------------------------------------------------------------------
%% 
\pts{10}
\begin{problem}
Design the \texttt{merge} method of the the \texttt{MergeAlgorithm}
class, whose purpose is to take two sorted lists (both are assumed to
be ordered by a given comparator) and to produce a single merged list
that contains all the elements of the two lists in sorted order.

\begin{verbatim}
import java.util.*;
import tester.*;

class MergeAlgorithm {

  <T> ArrayList<T> merge(ArrayList<T> list1,
                         ArrayList<T> list2,
                         Comparator<T> comp) {
         ...
  }
}
\end{verbatim}

\noindent
The \texttt{merge} method should not mutate any of its arguments.

\begsol{\vspace{0.5in}}
\begin{verbatim}
class Merge {

    // Merge two sorted lists into a single sorted list.
    // ASSUME: list1, list2 are sorted according to comp.
    <T> ArrayList<T> merge(ArrayList<T> list1,
			   ArrayList<T> list2,
			   Comparator<T> comp) {

	ArrayList<T> result = new ArrayList<T>();
	Integer i1 = 0;
	Integer i2 = 0;

	while (!i1.equals(list1.size()) &&
	       !i2.equals(list2.size())) {

	    if (comp.compare(list1.get(i1), list2.get(i2)) < 0) {
		result.add(list1.get(i1));
		i1 = i1 + 1;
	    } else {
		result.add(list2.get(i2));
		i2 = i2 + 1;
	    }
	}
	while (!i1.equals(list1.size())) {
	    result.add(list1.get(i1));
	    i1 = i1 + 1;
	}
	while (!i2.equals(list2.size())) {
	    result.add(list2.get(i2));
	    i2 = i2 + 1;
	}

	return result;
    }
}

class CharComp implements Comparator<Character>{
  public int compare(Character c1, Character c2){
    return c1.compareTo(c2);
  }
}

class ExamplesMerge{
  ExamplesMerge(){}

  ArrayList<Character> alphabet1 = 
      new ArrayList<Character>(Arrays.asList(
          'a', 'c', 'e', 'g', 'i', 
          'k', 'l', 'm', 'n', 'o', 'p', 'q', 'r', 's'));

  ArrayList<Character> alphabet2 = 
      new ArrayList<Character>(Arrays.asList(
          'b', 'd', 'f', 'h', 'j', 
          't', 'u', 'v', 'w', 'x', 'y', 'z'));

  ArrayList<Character> alphabet = 
      new ArrayList<Character>(Arrays.asList(
          'a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j', 
          'k', 'l', 'm', 'n', 'o', 'p', 'q', 'r', 's', 
          't', 'u', 'v', 'w', 'x', 'y', 'z'));

  ArrayList<Character> result = 
      new ArrayList<Character>();

  CharComp comp = new CharComp();
  Merge ma = new Merge();

  void testMerge(Tester t){
    t.checkExpect(ma.merge(alphabet1, alphabet2, comp, result), alphabet);
  }
}
\end{verbatim}
\endsol

\end{problem}

\newpage
[Here is some more space for the previous problem.]

\newpage
%% -----------------------------------------------------------------------------
%% 
\pts{15}
\begin{problem}
Class \texttt{CircularList} is similar to the \texttt{Dequeue} you
have designed in your homeworks, but has only one head node and the
last node in the list is connected back to the head - making the list
circular.

\begin{verbatim}
// to represent one node in a circular list
class Node<T>{
  T data;
  Node<T> prev;
  Node<T> next;

  Node(T data, Node<T> prev, Node<T> next){
    this.data = data;
    this.prev = prev;
    this.next = next;
  }

}

// to represent a circular list
class Circular<T>{
  Node<T> ht;

  Circular(){
    this.ht = new Node(null, null, null);
    this.ht.prev = this.ht;
    this.ht.next = this.ht;
  }
...
}
\end{verbatim}

Design the following methods for this circular list:

\begin{enumerate}
\newpage
\item Design the method \texttt{isEmpty()} for the
  \texttt{Circularlist}.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  // is this list empty?
  public boolean isEmpty(){
    return(this.ht.next == this.ht);
  }

\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{void addAtFront(T t)} for the
  \texttt{Circularlist} that adds the node with given item as its
  value at the front of this list.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  // add the given item to the front of this list
  public void addAtFront(T t){
    Node<T> oldLast = this.ht.next;
    Node<T> n = new Node<T>(t, this.ht, oldLast);
    this.ht.next = n;
    oldLast.prev = n;
  }
\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{void addAtTail(T t)} for the
  \texttt{Circularlist} that adds the node with given item as its
  value at the tail of this list.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  //add the given item to the tail of this list
  public void addAtTail(T t){
    Node<T> oldLast = this.ht.prev;
    Node<T> n = new Node<T>(t, oldLast, this.ht);
    this.ht.prev = n;
    oldLast.next = n;
  }
\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{ T removeFromFront()} for the
  \texttt{Circularlist} that removes the node at the front and returns
  its value.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  //remove the given item from the front of this list
  public T removeFromFront(){
    if (this.isEmpty())
      throw new RuntimeException("Cannor remove from an empty list");
    else{
      Node<T> removed = ht.next;
      ht.next = removed.next;
      removed.next.prev = ht;
      return removed.data;
    }
  }
\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{ T removeFromTail()} for the
  \texttt{Circularlist} that removes the node at the tail and returns
  its value.

\begsol{\vspace{0.5in}}
\begin{verbatim}
  //remove the given item from the tail of this list
  public T removeFromTail(){
    if (this.isEmpty())
      throw new RuntimeException("Cannot remove from an empty list");
    else{
      Node<T> removed = ht.prev;
      ht.prev = removed.prev;
      removed.prev.next = ht;
      return removed.data;
    }
  }
\end{verbatim}
\endsol

\newpage
[Here is some more space for the previous problem.]

\begsol{\vspace{0.5in}}
\begin{verbatim}
class ExamplesCircular{
  ExamplesCircular(){}

  Circular<String> cir = new Circular<String>();

  void testCircular(Tester t){
    t.checkExpect(this.cir.isEmpty(), true);
    //cir.addAtFront("hello");
    //cir.addAtFront("ciao");
    //cir.addAtFront("bye");

    cir.addAtTail("hello");
    cir.addAtTail("ciao");
    cir.addAtTail("bye");

    System.out.println("Adding at the tail:");

    System.out.println(this.cir.ht.next.data);
    System.out.println(this.cir.ht.next.next.data);
    System.out.println(this.cir.ht.next.next.next.data);

    System.out.println(this.cir.ht.prev.data);
    System.out.println(this.cir.ht.prev.prev.data);
    System.out.println(this.cir.ht.prev.prev.prev.data);

    t.checkExpect(this.cir.isEmpty(), false);
    t.checkExpect(this.cir.removeFromTail(), "bye");
    t.checkExpect(this.cir.removeFromTail(), "ciao");
    t.checkExpect(this.cir.removeFromFront(), "hello");

    t.checkExpect(this.cir.isEmpty(), true);

    cir.addAtFront("hello");
    cir.addAtFront("ciao");
    cir.addAtFront("bye");

    System.out.println("Adding at the front:");

    System.out.println(this.cir.ht.next.data);
    System.out.println(this.cir.ht.next.next.data);
    System.out.println(this.cir.ht.next.next.next.data);

    System.out.println(this.cir.ht.prev.data);
    System.out.println(this.cir.ht.prev.prev.data);
    System.out.println(this.cir.ht.prev.prev.prev.data);

    t.checkExpect(this.cir.isEmpty(), false);
    t.checkExpect(this.cir.removeFromFront(), "bye");
    t.checkExpect(this.cir.removeFromFront(), "ciao");
    t.checkExpect(this.cir.removeFromTail(), "hello");

    t.checkExpect(this.cir.isEmpty(), true);
  }
}
\end{verbatim}
\endsol

\end{enumerate}
\end{problem}

\newpage
%% -----------------------------------------------------------------------------
%% 
\pts{10}
\begin{problem}
You are now masters of encoding and decoding algorithms. Here is a new
one for you to implement: When encoding the letters, leave the first
letter the same as the original. Replace the second letter of the
message with a letter that is one space later in the alphabet. (So, if
the second letter was `e', you would replace it with `f'.) Replace the
third letter with one that is two letters later in the alphabet. So,
if the third letter was `l', you would replace it with `n'.) Proceed
like this till the end. Of course, circle around to the beginning of
the alphabet, if you get to the end. (So, if `w' appears as the 7th
letter, it will be replaced by `b' --- we included a 27th character,
the space.)

For example, ``hello world'' would be encoded as ``hfnosebvzun''.

We are giving you the \texttt{ArrayList<Character>} that contains all
letters of the alphabet and the space character at the end:

\begin{verbatim}
interface DTree {
  /** The original list of characters used in encoding and decoding */
  ArrayList<Character> alphabet = 
        new ArrayList<Character>(Arrays.asList(
            'a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j', 
            'k', 'l', 'm', 'n', 'o', 'p', 'q', 'r', 's', 
            't', 'u', 'v', 'w', 'x', 'y', 'z', ' '));
}
\end{verbatim}

\newpage
\begin{enumerate}
\item Design the method \texttt{encode} that consumes a message
  \texttt{String} and produces the encoded \texttt{String}

\begsol{\vspace{0.5in}}
\begin{verbatim}
class Encoder implements DTree{
  Encoder(){}

  String encode(String source, String acc){
    if (source.length() > 0)
    return encode(source.substring(1),
        acc + alphabet.get((alphabet.indexOf(source.charAt(0)) + acc.length()) % 27));
    else 
      return acc;
  }

  String decode(String source, String acc){
    if (source.length() > 0)
    return decode(source.substring(1),
        acc + alphabet.get((alphabet.indexOf(source.charAt(0)) - acc.length()+ 27) % 27));
    else 
      return acc;
  }

}
\end{verbatim}
\endsol

\newpage
\item Design the method \texttt{decode} that consumes an encoded message
  \texttt{String} and produces the original message \texttt{String}

\begsol{\vspace{0.5in}}
\begin{verbatim}
class ExamplesExam3{
  ExamplesExam3(){}

  Encoder enc = new Encoder();

  void testEncoder(Tester t){
    t.checkExpect(enc.encode("hello world", ""), "hfnosebvzun");
    t.checkExpect(enc.encode("how are you", ""), "hpycewkgfxd");

    t.checkExpect(enc.decode("hfnosebvzun", ""), "hello world");
    t.checkExpect(enc.decode("hpycewkgfxd", ""), "how are you");
  }

  public static void main(String[] argv){
    ExamplesExam3 e = new ExamplesExam3();
    Encoder enc = new Encoder();

    System.out.println("hello world" + " encoded as " + 
        enc.encode("hello world", ""));
    System.out.println("how are you" + " encoded as " + 
        enc.encode("how are you", ""));

    System.out.println("hello world" + " decoded as " + 
        enc.decode("hfnosebvzun", ""));
    System.out.println("how are you" + " decoded as " + 
        enc.decode("hpycewkgfxd", ""));

    Tester.runFullReport(e);
  }
}
\end{verbatim}
\endsol

\end{enumerate}
\end{problem}

\end{document}