### 2013-09-29 tfpie/response.txt
review 3

----------------------- REVIEW 1 ---------------------

I use the term "work-in-progress," because the authors have
side-stepped the methodology used to introduce types and the
transition from their tower of student languages to an industrial
language. Granted, this is likely something the authors are still
polishing. Nonetheless, it is important to provide a vision for how to
do these things and to outline what they have done highlighting both
succeses and failures. In this manner, interested readers may join
their efforts by reproducing the authors' methodology in their schools
using the languages developed by the authors. Let's face it,
introducing types is necessary to make the move to an
industrial-strength language and without providing a vision for it
many may balk. I suspect that the transition builds on the informal
contracts students are familiar with. Thus, adding this vision should
not be to difficult. Outlining this vision may obviate doubts readers
may have about making the transition to an industrial-strength
language. This is the primary reason that I am unable to put this
article in the strong accept column.

RESPONSE: We have added a paragraph in section 4.1 on this subject.

Page 2

Spell out an example of the details of numeric representation that
crop up quickly.

RESPONSE: We have added an example.

Implicit in the discussion is the assumption that reaching industrial
language competence by the end of the semester is a desirable
goal. The authors should address this head on. An obvious alternative
is emphasize and reinforce OO concepts and leave the industrial
language for latter (one semester?) in the curriculum. I am aware that
at NE students are expected to get a co-op/intership over the
summer. That, however, is a NE pecularity and is not a pedagogical
argument.

RESPONSE: We make explicit in section 2 the aspects of Northeastern
that inform our course.  It is not only co-op, but downstream courses,
that require Java.

Page 3

"sent by using the send keyword, followed by an object, a message
name, and some number of arguments"

        What are these "some number of arguments?"
        They are not part of the examples.

RESPONSE: These particular methods take no arguments, but the `dist`
method, on the same page, takes an argument.

Page 6

Spell out for the reader how students are explained the world
parameter is not needed and, therefore, removed from the methods in
the class definition.

RESPONSE: This is already covered in the discussion of OO
fundamentals.

In the class version, how does the runtime know which method to use?
Are the names of the handlers restricted to on-tick, on-key, etc.? Do
the methods have to appear in a certain order? Does the mechanism used
to make the association of events with the proper handler pose any
difficulty for the students? One nice feature of the functional
approach is that such associations are made explicitly and this allows
students to name their handlers whatever makes sense to them.

RESPONSE: We made this more explicit in the paper.

Do you mean the first interactive event-driven program?

RESPONSE: No, this is the first program we show.

The programs in Figure 2 do not animate a rocket. Please reformulate
the description of the programs in Figure 2.

RESPONSE: We have revised the program.

Show the reader how you use with students in class the state pattern
to animate a rocket that lands. There is plenty of room within the
20-page limit.

RESPONSE: We added an appendix with this program.

Page 11

Spell out the drawbacks that are alleviated and those that are not
addressed by DrJava and BlueJ. As a reader, I want to clearly see the
contrast of your approach with these.

RESPONSE: We now make this explicit.

Page 12

Can you quantify the recruitment/retention impact on female students?
Give the reader a sense of what you mean. It must be significant given
that you mention it!

RESPONSE: A quantitative evaluation would be inappropriate in this
situation. We have clarified the point.

----------------------- REVIEW 2 ---------------------

p2 "a set that is demographically skewed" - in what way?  Also, is
this first paragraph written from your own experience of past versions
of the course?

RESPONSE: We commented on our prior experience.

p5 I think this is the first time the names "BSL" and "Class" are
mentioned.  As far as I know the name BSL is taken from HTDP, and
there does seem to be a few places where it is assumed the reader is
familiar with the content of HTDP. Maybe it would be good to give a
brief summary of the HTDP approach and what these languages are
earlier.

RESPONSE: We clarify what these languages are.

p7 Is this the first mention of "Intermediate Student" language?
Again, maybe in section two where the idea of language levels is
mentioned, maybe it would be good to go into a bit more detail for the
benefit of those who haven't read HTDP.

RESPONSE: We extended section 2 to cover more HTDP background.

p9 Is it possible to include the Racket version alongside this
program? I found it interesting to compare them, and there does seem
to be enough space on the page for it.

RESPONSE: The BSL and Class versions of this program are in figure 1.

p10 "functional OO programming" reads a bit strangely to me, since it
could mean "functional" in a non-technical sense. Maybe "a functional
subset of ..."?

RESPONSE: Clarified.

p12 It's an interesting comment about retaining female students, but
it's not clear how the design of the course might be relevant to
this. Do you have any suggestions why? Could it just be a random
fluctuation?

RESPONSE: We clarified the anecdotal nature of this.

----------------------- REVIEW 3 ---------------------

(1) in the introduction, page 1, the bullet-list enumerates
'significant transitions happen in concert'. In the paper we learn how
to overcome bullet 1, but not bullets 2, 3, 4, and 5.

RESPONSE: Our approach presents OO programming in a context that
avoids _all 5_ of these significant transitions until after the
fundamental concepts of OO programming are taught, as we emphasize.

On page 8, first
paragraph: "In particular: objects, classes, fields and methods,
dynamic dispatch, inheritance, and overriding." are enumerated as
being essential concepts of object-oriented programming that students
have had experience with in the new set-up. The paper explains this
for the first half ('objects, classes, fields and methods') but not
for the second half ('dynamic dispatch, inheritance, and overriding').

Suggestion for improvement: extend section 3 with material how dynamic
dispatch, inheritance, and overriding is presented to students.

RESPONSE: We discuss dynamic dispatch extensively in the paper.  This
is the entire point of section 3.2.  We added a paragraph discussion
inheritance and overriding.

(2) Section 5.1 is too long to my taste: a lot of the criticism has
been discussed in the motivation of the introduction of this paper.

Suggestion for improvement: shorten section 5.1. The extra space can
be used to elaborate on issue (1) above. I don't think this section
requires sub sections, so all text can be organized into a single
section.

RESPONSE: We feel that this discussion is needed.

(3) In the introduction, it is said that 'an alternative approach to
the second semester [that] overcomes these problems and provides a
gradual introduction to OO programming'. However, acting as the
devil's advocate, I can argue that this is not true: halfway semester
2 there is still the abrupt switch to both another language, Java,
*and* another tool set, Eclipse. Stated bluntly, the switch is only
postponed instead of solved.

RESPONSE: We clarify how the transition is less troublesome here.  We
also describe future work that would further alleviate these
difficulties.

"numerous observers [9, 23]": only two references, of which [23] is
not scientific is not particularly 'numerous'. Please provide more
scientific references or remove the word 'numerous' (I prefer the
first option)

RESPONSE: We have added additional references.

----------------------- REVIEW 4 ---------------------
PAPER: 1
TITLE: From Principles to Practice with Class in the First Year
AUTHORS: Sam Tobin-Hochstadt and David Van Horn

OVERALL EVALUATION: 2 (accept)
REVIEWER'S CONFIDENCE: 4 (high)

----------- REVIEW -----------

The authors describe a new approach to tackle difficulties in
transitioning from a functional semester one course to an
object-oriented semester two course. Students are reportedly confused
by the circumstantial differences (compilation model / environment /
syntax) which makes grasping the fundamental concepts being taught
harder to grasp. Instead of transitioning on day one of the second
semester to Java in Eclipse (or similar), the authors present an
approach that continues the progression of student languages to
include objects and other OO concepts before transitioning to a "real"
OO environment and language.

The paper is well-written, is well-positioned in the context of the
status quo and balances the pedagogical approach with the material
used. The idea of language levels is properly motivated and the
continuation of this approach beyond the first semester seems - after
reading the paper - the Obvious Choice. What is lacking, however, is a
motivation for the point at which the transition to Java is made. Why
aren't e.g. iteration, mutability & typing introduced in the similar
vein as objects, class-hierarchies, etc.? Was this a choice from
practicality, or is there a reason to introduce them post-transition?

A few specific comments and minor corrections:

 - (Sec.1) "common approache to" --> "common approach to"

 - (Sec.1) "interaction-oriented language" / "batch-oriented language"
   --> it seems the authors are refering to the environments/tooling
   rather than the language. If not, this terminology warrents
   clarification.

 - (Sec.1) "aspects of course remain" --> "aspects of *the* course
   remain"

 - (Sec.2) "while most students do some software development, there is
   almost no uniformity beyond that" & "in which they will use
   professional-grade languages and tools which will almost certainly
   be object-oriented" are, at least, paradoxical (if not
   contradictory). This conflict in the motivation should be addressed
   for the final version of the paper.

 - (Sec.3.1) "and message are sent" --> "and message*s* are sent"

 - (Sec.4.2) "Thanks to the the preparation" --> omit one "the"

- (Sec.6) "The class has also had notable success in the recruitment
  and retention of female students, as compared to the other versions
  of the second-semester course." --> I find this a rather hefty
  statement. It would - in my mind - at least warrent further
  research. If this consistently holds true, there are valuable
  insights to be gained about pedagogical gender-script in the field
  of CS. The statement by itself seems a bit hefty to be included
  without at least some concrete data and/or qualitative evaluation.