### 2018-11-05 1/exams/m2.tex
Midterm 2.

\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{xcolor,censor}
\usepackage{fancyvrb}
\usepackage{fontspec}
\usepackage{alltt}

\setmainfont{Times New Roman}

\censorruleheight=0ex
%\StopCensoring

\begin{document}

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 2}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2018 \ \\
  }
\end{center}

\vspace{2em}

\noindent
NAME:\verb|________________________________________________________|

\vspace{2em}

\noindent
UID: \verb|__________________________________________|

\vspace{2em}

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 15 \\ \hline
  3 & 15 \\ \hline
  4 & 20 \\ \hline
  5 & 20 \\ \hline\hline
  Total: & 80\\
  \hline
\end{tabular}
\end{center}

\noindent
This test is open-book, open-notes, but you may not use any computing
device other than your brain and may not communicate with anyone.
You have 50 minutes to complete the test.

\vskip 1em

\noindent
The phrase ``design a program'' or ``design a function'' means follow
the steps of the design recipe.  Unless specifically asked for, you do
not need to provide intermediate products like templates or stubs,
though they may be useful to help you construct correct solutions.

\vskip 1em

\noindent
You may use any of the data definitions given to you within this exam
and do not need to repeat their definitions.

\vskip 1em

\noindent
Unless specifically instructed otherwise, you may use any built-in ISL+
functions or data types.

\vskip 1em

\noindent
When writing tests, you may use a shorthand for writing check-expects
by drawing an arrow between two expressions to mean you expect the
first to evaluate to same result as the second.  For example, you may
write \verb|(add1 3)| $\rightarrow$ \verb|4| instead of
\verb|(check-expect (add1 3) 4)|.

\newpage

\noindent
{\bf Problem 1 (10 points).}
%
Here is a data definition for representing definitions in a
dictionary:
\begin{verbatim}
;; A Defn is a (make-defn String String)
(define-struct defn (word meaning))
;; Interp: a word and its definition
;; Example:
(make-defn "pair" 
           "a set of two things used together or regarded as a unit")
\end{verbatim}

Design the following function for determining whether one definition
should come before another in a dictionary.  (Hint: you can use the
function {\tt string<?} to help.)
\begin{verbatim}
;; defn<? : Defn Defn -> Boolean
;; Does the word defined in d1 come before d2, alphabetically?
(define (defn<? d1 d2) ...)
\end{verbatim}

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

(check-expect (defn<? (make-defn "a" "first") (make-defn "b" "second")) #true)
(check-expect (defn<? (make-defn "b" "second") (make-defn "a" "first")) #false)
(define (defn<? d1 d2)
  (string<? (defn-word d1) (defn-word d2)))

\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 2 (15 points).}
%
A dictionary can be represented using a structure similar to a binary
search tree as follows (where {\tt Defn} is defined in problem 1):
\begin{verbatim}
;; A BSD (binary search dictionary) is one of:
;; - (make-leaf)
;; - (make-node d l r), where d is a Defn, l is a BSD, r is a BSD,
;;      and every definition in l alphabetically comes before d
;;      and d comes alphabetically before every definition in r
;; ASSUME: words are defined at most once in a BSD
(define-struct leaf ())
(define-struct node (val left right))
\end{verbatim}

Design a function {\tt all-words} that consumes a {\tt BSD} and
produces a list of all the words (just the words, not their meanings)
defined in the dictionary, in alphabetic order.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

(define a (make-defn "a" "first"))
(define b (make-defn "b" "second"))
(define c (make-defn "c" "third"))
(define l (make-leaf))

;; all-words : BSD -> [Listof String]
;; List all words defined in given dictionary in alphabetic order
(check-expect (all-words l) '())
(check-expect (all-words (make-node b (make-node a l l) (make-node c l l)))
              (list "a" "b" "c"))
(define (all-words bsd)
  (cond [(leaf? bsd) '()]
        [(node? bsd)
         (append (all-words (node-left bsd))
                 (list (defn-word (node-val bsd)))
                 (all-words (node-right bsd)))]))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 3 (15 points).}
%
For each of the following functions, correctly reformulate the
function in terms of an existing abstraction function.  You may use
either lambda-expressions or local definitions if needed.

\begin{alltt}
;; any-zero? : [Listof Number] -> Boolean
;; Are any of the numbers in given list equal to zero?
(check-expect (any-zero? '(1 2 3)) #false)
(check-expect (any-zero? '(1 0 3)) #true)
(define (any-zero? xs)
  (cond [(empty? xs) #false]
        [(cons? xs)
         (or (zero? (first xs))
             (any-zero? (rest xs)))]))
\end{alltt}

\vspace{.8in}

\begin{SaveVerbatim}{VerbEnv}
SOLUTION:

(define (any-zero? xs) (ormap zero? xs))

\end{SaveVerbatim}

\censor{%
\vspace{-.8in}
\noindent
\BUseVerbatim{VerbEnv}}

\begin{alltt}
(define DOT (circle 10 "solid" "black"))
(define MT (empty-scene 200 200))

;; draw-dots : [Listof Posn] -> Image
;; Draw dots for all given positions on an empty scene.
(check-expect (draw-dots '()) MT)
(check-expect (draw-dots (list (make-posn 10 20) (make-posn 20 30)))
              (place-image DOT 10 20 (place-image DOT 20 30 MT)))
(define (draw-dots ps)
  (cond [(empty? ps) MT]
        [(cons? ps)
         (place-image DOT
                      (posn-x (first ps))
                      (posn-y (first ps))
                      (draw-dots (rest ps)))]))
\end{alltt}

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

(define (draw-dots ps)
  (foldr (lambda (p img) (place-image DOT (posn-x p) (posn-y p) img))
         (empty-scene 200 200)
         ps))

\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage
\noindent
{\bf Problem 3 (cont).}

\begin{alltt}
;; remove-short-words : Number [Listof Defn] -> [Listof Defn]
;; Remove all definitions for words shorter than given length
(check-expect (remove-short-words 2 (list (make-defn "a" "first") 
                                          (make-defn "ab" "rest")))
              (list (make-defn "ab" "rest")))
(define (remove-short-words len ds)
  (cond [(empty? ds) '()]
        [(cons? ds)
         (if (< (string-length (defn-word (first ds))) len)
             (remove-short-words len (rest ds))
             (cons (first ds)
                   (remove-short-words len (rest ds))))]))
\end{alltt}

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

(define (remove-short-words len ds)
  (filter (lambda (d) (not (< (string-length (defn-word d)) len))) ds))

\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage

\noindent
{\bf Problem 4 (20 points).}
%
Design an abstraction of the following two functions and re-create
the original functions in terms of your abstraction.

\begin{verbatim}
;; sqr-pos : [Listof Number] -> [Listof Number]
;; Square each number in the list that is positive, remove the rest
(check-expect (sqr-pos '(2 -3 4)) '(4 16))
(define (sqr-pos xs)
  (cond [(empty? xs) '()]
        [(cons? xs)
         (if (positive? (first xs))
             (cons (sqr (first xs))
                   (sqr-pos (rest xs)))
             (sqr-pos (rest xs)))]))

;; lowercase-lengths : [Listof String] -> [Listof Number]
;; Produce a list of lengths for all lowercase strings in given list
(check-expect (lowercase-lengths '("Hey" "you" "guys")) '(3 4))
(define (lowercase-lengths xs)
  (cond [(empty? xs) '()]
        [(cons? xs)
         (if (string-lower-case? (first xs))
             (cons (string-length (first xs))
                   (lowercase-lengths (rest xs)))
             (lowercase-lengths (rest xs)))]))
\end{verbatim}

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:
;; filter&map : [Listof X] [X -> Boolean] [X -> Y] -> [Listof Y]
(define (filter&map xs p f)
  (cond [(empty? xs) '()]
        [(cons? xs)
         (if (p (first xs))
             (cons (f (first xs))
                   (filter&map (rest xs) p f))
             (filter&map (rest xs) p f))]))

(define (sqr-pos xs)
  (filter&map xs positive? sqr))

(define (lowercase-lengths xs)
  (filter&map xs string-lower-case? string-length))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage

\noindent
[Space for problem 4.]

\newpage

\noindent
{\bf Problem 5 (20 points).}
%
Consider the following data definition (and examples) for representing sentences:
\begin{verbatim}
;; A Sentence is a [NEListof String]

(define sent1 (list "a" "dog" "barked"))
(define sent2 (list "a" "cat" "meowed"))
\end{verbatim}
The {\tt [NEListof X]} data definition is used to represent 
non-empty lists of elements:
\begin{verbatim}
;; A [NEListof X] is one of:
;; - (cons X '())
;; - (cons X [NEListof X])
\end{verbatim}
Given a {\tt Sentence}, a \emph{bigram} is a list of two consecutive
elements of the sentence. For example, {\tt sent1} has exactly two
bigrams:
\begin{verbatim}
(list "a" "dog")
(list "dog" "barked")
\end{verbatim}
Using the following data definition,
\begin{verbatim}
;; a Bigram is a (list String String)
\end{verbatim}
design a function {\tt bigrams} which takes a {\tt Sentence} and
returns a list of {\tt Bigrams}.

\begin{SaveVerbatim}{VerbEnv}

SOLUTION:

;; bigrams : Sentence -> [Listof Bigram]
;; List of all adjacent words in a sentence
(check-expect (bigrams (list "one")) '())
(check-expect (bigrams sent1)
              (list (list "a" "dog") (list "dog" "barked")))
(define (bigrams sent)
  (cond [(empty? (rest sent)) '()]
        [else
         (cons (list (first sent) (second sent))
               (bigrams (rest sent)))]))
\end{SaveVerbatim}

\censor{%
\noindent
\BUseVerbatim{VerbEnv}}

\newpage
\noindent
[Space for problem 5.]

\end{document}